---
title: "fire-color: preview Data"
subtitle: "identifying post-fire color changes in lakes of the western U.S."
author: "Bella Oleksy et al. "
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    latex_engine: pdflatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir='..')
knitr::opts_chunk$set(out.width = '100%',fig.height=8,
                      echo=FALSE, 
               warning=FALSE, message=FALSE) 

```



```{r some package version control stuff, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
##Hi! If you have never used renv read their little vignette here. 
##https://rstudio.github.io/renv/articles/renv.html

##If you've ever had script not work after updating R/various libraries, using
##renv() is a nice way to avoid a lot of future headaches. -IAO

if (!require('renv')) install.packages('renv');library('renv')

#Create a lockfile for packages
# renv::init()
```


```{r Load necessary packages, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library('here')

##enable easy file referencing in project-oriented workflows.
##https://rstats.wtf/project-oriented-workflow.html
## ^ ^ a little more on that

source("fire-color/scripts/00_libraries.R")



```


```{r Load LAGOS IDs dataframe, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
source("fire-color/scripts/01.0_pullLakeIDsLAGOS.R")
source("fire-color/scripts/01.1_pullLakeCat.R")
#Currently only pulling out MTBS, FirePerimeters, and ForestLoss, but all the other data is in that R file and commented out for speed. 

```


```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#Describes proportion of catchment / watershed within burn perimeters
#But I think I'm missing the burn severity numbers 
MTBS_long <- lakeCat %>%
  dplyr::select(nhdplusv2_comid, lake_lon_decdeg, lake_lat_decdeg, contains("MTBS"))%>%
  pivot_longer(-(1:3), names_pattern = "(\\d+)([A-Za-z]+)", names_to = c("year","scale"))
    # "([A-Za-z]+)" matches only letters, so will pull out just "Ws" or "Cat". "(\\d+)" matches digits and will pull out the year.
  
names(MTBS_long)
```

# MTBS
## Histograms of % area burned by Ws and Cat scales

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = '100%', fig.height=8}
MTBS_long %>%
  filter(value>1)%>%
  filter(scale=="Ws") %>%
  mutate(year=as.numeric(year))%>%
  ggplot(aes(value, group=year)) +
  geom_histogram(aes(fill=year)) +  # scale histogram y
  # geom_density(aes(color=year), alpha=0.4) +
  facet_wrap(~year, scales="free") +
  scale_fill_continuous_sequential(palette="ag_Sunset")+
  scale_color_continuous_sequential(palette="ag_Sunset")+
  ggtitle("MTBS % area burned - Ws scale")+
  ggpubr::theme_pubr(base_size=8)
```
\newpage
```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = '100%', fig.height=8}
MTBS_long %>%
  filter(value>1)%>%
  filter(scale=="Cat") %>%
  mutate(year=as.numeric(year))%>%
  ggplot(aes(value, group=year)) +
  geom_histogram(aes(fill=year)) +  # scale histogram y
  # geom_density(aes(color=year), alpha=0.4) +
  facet_wrap(~year, scales="free") +
  scale_fill_continuous_sequential(palette="ag_Sunset")+
  scale_color_continuous_sequential(palette="ag_Sunset")+
  ggtitle("MTBS % area burned - Cat scale (local)") +
  ggpubr::theme_pubr(base_size=8)
```

In every year there are always a number of lakes that burn severly (>90% of the total area of every local catchment. If you look at the full watershed (Ws) scale there are a lot more lakes on the low end (<20%).

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = '100%', fig.height=8}

#Number of lakes that experienced forest loss due to fire within local catchment
MTBS_summary<- left_join(MTBS_long%>%
  filter(value>1)%>%
  filter(scale=="Cat") %>%
  group_by(year) %>%
  summarize(n_lakes_Cat=length(unique(nhdplusv2_comid))) ,

#Number of lakes that experienced forest loss due to fire 2000-2010 within larger watershed
MTBS_long%>%
  filter(value>1)%>%
  filter(scale=="Ws") %>%
  group_by(year) %>%
  summarize(n_lakes_Ws=length(unique(nhdplusv2_comid))) ,

by="year")


knitr::kable(MTBS_summary, "simple", caption="Table summarizing the number of lakes with any percentage burn (>1%) at the local catchment (Cat) and full watershed (Ws) scale, by year")
```

Looks promising, we have hundreds of lakes lakes that were close to forests fires in every year and the numbers fluctuate quite a bit year-to-year.

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = '100%', out.height='100%'}
MTBS_summary %>%
  pivot_longer(-1)%>%
  ggplot(aes(x=year,y=value, fill=name, color=name, group=name))+
  geom_line()+
  geom_point(shape=21, size=2.5)+
  scale_color_manual(values=c("#90a955","#31572c"),
                    labels=c("Catchment","Watershed"),
                    name="Legend")+
  scale_fill_manual(values=c("#90a955","#31572c"),
                    labels=c("Catchment","Watershed"),
                    name="Legend")+
  labs(x="Year",
       y="n lakes",
       title="Number of lakes in burned watersheds/local catchments")+
  ggpubr::theme_pubr(base_size=8)+
  theme(axis.text.x=element_text(angle = 45, hjust = 1, size=8))


```

Looks like there is a step-change increase in the number of lakes in burned watersheds in the early 2000s, though no breakpoint detected using segmented regerssion. 

```{r, include=TRUE, error=FALSE}
# -------------------
# analyse breakpoints
# -------------------
# http://cran.r-project.org/doc/Rnews/Rnews_2008-1.pdf
if (!require('segmented')) install.packages('segmented');library('segmented')

MTBS_summary2<- MTBS_summary %>%
  mutate(year=as.numeric(as.character(year)))

# create a linear model
my.lm <- lm(n_lakes_Cat ~ year, data = MTBS_summary2)
summary(my.lm)

# # have to provide estimates for breakpoints.
# # after looking a the data, 
# my.seg <- segmented(my.lm, 
#                     seg.Z = ~ year, 
#                     psi = NA)
# 
# # When not providing estimates for the breakpoints "psi = NA" can be used.
# # The number of breakpoints that will show up is not defined
# #my.seg <- segmented(my.lm, 
# #                    seg.Z = ~ DistanceMeters, 
# #                    psi = NA)
# 
# # display the summary
# summary(my.seg)

```
\newpage
## Water color proof of concept

How many lakes do we have where the local catchment burned >90%? We think that is where we are most likely to see a color response if there is one. 

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = '100%', fig.height=5}

#Number of lakes that experienced forest loss due to fire within local catchment
MTBS_summary_severe<- MTBS_long%>%
  filter(value>90)%>%
  filter(scale=="Cat") %>%
  group_by(year) %>%
  summarize(n_lakes_Cat=length(unique(nhdplusv2_comid)))


knitr::kable(MTBS_summary_severe, "simple", caption="Table summarizing the number of lakes with any percentage burn (>90%) at the local catchment (Cat) and full watershed (Ws) scale, by year")
```

We have a total of `r MTBS_summary_severe %>% summarize(n_lakes=sum(n_lakes_Cat)) %>% pull()` lakes to check out. 

```{r, echo=FALSE, message=FALSE, warning=FALSE,out.width = '100%', out.height='100%'}
MTBS_summary_severe %>%
  ggplot(aes(x=year,y=n_lakes_Cat, group=NA))+
    geom_line()+
  geom_point(shape=21, size=2.5, fill="#90a955")+
  labs(x="Year",
       y="n lakes",
       title="Number of lakes where >90% of the local catchment burned")+
  ggpubr::theme_pubr(base_size=12)+
  theme(axis.text.x=element_text(angle = 45, hjust = 1, size=8))


```

## Write .csv file with comids for all lakes with fires
```{r, echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}

# MTBS_export<-MTBS_long %>%
#   filter(value>0)%>%
#   dplyr::select(nhdplusv2_comid, lake_lon_decdeg, lake_lat_decdeg)
# write_csv(MTBS_export, "data_export/MTBS_burnedLakes.csv")

```

<!-- #Does the number of small fires look like it's increasing? -->
<!-- MTBS_long %>% -->
<!--   filter(value>0 & value<20)%>% -->
<!--   filter(scale=="Cat") %>% -->
<!--   mutate(year=as.numeric(year))%>% -->
<!--   ggplot(aes(value, group=year)) + -->
<!--   scale_color_continuous_sequential(palette="ag_Sunset")+ -->
<!--   geom_density(aes(color=year)) + -->
<!--   # facet_wrap(~year) + -->
<!--   ggtitle("MTBS % area burned - Cat scale (local)")+ -->
<!--   theme_dark() -->


<!-- #Does the number of moderate fires look like it's increasing? -->
<!-- MTBS_long %>% -->
<!--   filter(value>20 & value<80)%>% -->
<!--   filter(scale=="Cat") %>% -->
<!--   mutate(year=as.numeric(year))%>% -->
<!--   ggplot(aes(value, group=year)) + -->
<!--   scale_color_continuous_sequential(palette="ag_Sunset")+ -->
<!--   geom_density(aes(color=year)) + -->
<!--   # facet_wrap(~year) + -->
<!--   ggtitle("MTBS % area burned - Cat scale (local)")+ -->
<!--   theme_dark() -->



<!-- names(FirePerimeters) -->

<!-- FirePeriments_long<- FirePerimeters %>% -->
<!--       pivot_longer(-(1:6), names_pattern = "([A-Za-z]+)(\\d+)([A-Za-z]+)", names_to = c("units","year","scale")) -->

<!-- FirePeriments_long %>% -->
<!--   filter(value>0 & value <25)%>% -->
<!--   filter(scale=="Ws") %>% -->
<!--   ggplot(aes(value)) + -->
<!--   geom_histogram() + -->
<!--   facet_wrap(~year, scales="free") + -->
<!--   ggtitle("% Forest loss to fire - Ws scale") -->

<!-- FirePeriments_long %>% -->
<!--   filter(value>0 & value <25)%>% -->
<!--   filter(scale=="Cat") %>% -->
<!--   ggplot(aes(value)) + -->
<!--   geom_histogram() + -->
<!--   facet_wrap(~year, scales="free") + -->
<!--   ggtitle("% Forest loss to fire - local Cat scale") -->



<!-- #Number of lakes that experienced forest loss due to fire 2000-2010 within local catchment -->
<!-- FirePeriments_long%>% -->
<!--   filter(value>0)%>% -->
<!--   filter(scale=="Cat") %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(n_lakes=length(unique(nhdplusv2_comid))) -->


<!-- #Number of lakes that experienced forest loss due to fire 2000-2010 within larger watershed -->
<!-- FirePeriments_long%>% -->
<!--   filter(value>0)%>% -->
<!--   filter(scale=="Ws") %>% -->
<!--   group_by(year) %>% -->
<!--   summarize(n_lakes=length(unique(nhdplusv2_comid))) -->



<!-- ``` -->

<!-- Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. -->
