---
title: "fire-color data vis tool"
output:
  html_document:
    fig_width: 6
    fig_height: 4
author: "Bella Oleksy"
date: "11/17/2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir='..')
knitr::opts_chunk$set(fig.width=6,fig.height=4,
                      echo=FALSE, 
               warning=FALSE, message=FALSE) 

```


```{r Load necessary packages, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library('here')

##enable easy file referencing in project-oriented workflows.
##https://rstats.wtf/project-oriented-workflow.html
## ^ ^ a little more on that

source("fire-color/scripts/00_libraries.R")
# source("scripts/00_libraries.R")

# load packages
library(dplyr)
library(ggplot2)
library(rjson)
library(jsonlite)
library(leaflet)
library(RCurl)

```


```{r Load LAGOS IDs dataframe, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
source("fire-color/scripts/01.0_pullLakeIDsLAGOS.R")
source("fire-color/scripts/01.1_pullLakeCat.R")
# source("scripts/01.0_pullLakeIDsLAGOS.R")
# source("scripts/01.1_pullLakeCat.R")
#Currently only pulling out MTBS, FirePerimeters, and ForestLoss, but all the other data is in that R file and commented out for speed.

```

```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#Describes proportion of catchment / watershed within burn perimeters
#But I think I'm missing the burn severity numbers 
MTBS_long <- lakeCat %>%
  dplyr::select(nhdplusv2_comid,lake_nhdid,nhdplusv2_reachcode,lake_namegnis, lake_lon_decdeg, lake_lat_decdeg, contains("MTBS"))%>%
  pivot_longer(-(1:6), names_pattern = "(\\d+)([A-Za-z]+)", names_to = c("year","scale"))
    # "([A-Za-z]+)" matches only letters, so will pull out just "Ws" or "Cat". "(\\d+)" matches digits and will pull out the year.
  
names(MTBS_long)
```


<!-- ## Get data -->
<!-- ```{r} -->
<!-- base_url <- "https://data.colorado.gov/resource/j5pc-4t32.json?" -->
<!-- full_url <- paste0(base_url, "station_status=Active", -->
<!--             "&county=BOULDER") -->
<!-- water_data <- getURL(URLencode(full_url)) -->

<!-- # you can then pipe this -->
<!-- water_data_df <- fromJSON(water_data) %>% -->
<!--   flatten(recursive = TRUE) # remove the nested data frame -->

<!-- # turn columns to numeric and remove NA values -->
<!-- water_data_df <- water_data_df %>% -->
<!--   mutate_at(vars(amount, location.latitude, location.longitude), list(as.numeric)) %>% -->
<!--   filter(!is.na(location.latitude)) -->

<!-- str(water_data_df) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- # Specify custom icon -->
<!-- url <- "http://tinyurl.com/jeybtwj" -->
<!-- water <- makeIcon(url, url, 24, 24) -->

<!-- water_locations_map <- leaflet(water_data_df) %>% -->
<!--   addTiles() %>% -->
<!--   addCircleMarkers(lng = ~location.longitude, -->
<!--                    lat = ~location.latitude) %>% -->
<!--   addMarkers(lng = ~location.longitude, lat = ~location.latitude, icon=water, -->
<!--              popup = ~paste0(station_name, -->
<!--                            "<br/>Discharge: ", -->
<!--                            amount)) %>% -->
<!--   #Add both station name and discharge. <br/> is a line break in html. -->
<!--   addProviderTiles("Stamen.Terrain") -->

<!-- water_locations_map -->

<!-- pal <- colorFactor(c("navy", "red", "green"), -->
<!--                    domain = unique(water_data_df$station_type)) -->
<!-- unique_markers_map_2 <- leaflet(water_data_df) %>% -->
<!--   addProviderTiles("Stamen.Terrain") %>% -->
<!--   addCircleMarkers( -->
<!--     color = ~pal(station_type), -->
<!--     stroke = FALSE, fillOpacity = 0.5, -->
<!--     lng = ~location.longitude, lat = ~location.latitude, -->
<!--     label = ~as.character(station_type), -->
<!--     popup = ~paste0(station_name, -->
<!--                            "<br/>Discharge: ", -->
<!--                            amount) -->
<!--   ) -->
<!-- unique_markers_map_2 -->
<!-- ``` -->

## MTBS maps


The maps below summarize Monitoring Trends in Burn Severity (MTBS) data, extracted from [EPA's LakeCat dataset](https://www.epa.gov/national-aquatic-resource-surveys/lakecat-dataset-0). Before diving into this project, we want to get a sense for just how many lakes in the western U.S. are affected by fires every year. This particular dataset spans the years 1984-2018. 

I'd like to use these maps to understand a few spatiotemporal patterns

- Can we see how the location of fires has changed over time?
- Where have the most severe burns taken places?
- Which areas have the most small fires? large fires? medium fires?

### Maps - changes through time

You can use the leaflet below to look at clusters of large fires through time, where each point represents a lake where >1% of the local catchment burned. The color of the point corresponds to the year of the burn (less than 5 lakes in the dataset burned more than once, so we excluded those for now), with darker tones corresponding to older burns. 

If you hover over the point, the year is displayed. If you click on a point, you will see the lake name, comid, and percentage of the local catchment that burned that year. 

Right away you can see some clustering. For example, the 1988 Yellowstone fire complex shows up pretty vividly in the northwestern corner of Wyoming. There appears to be a lot more redish/purple hues in California where we know wildfires have been increasing. 

```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.width=8, fig.height=7}

MTBS_mapdf1<-MTBS_long %>%
filter(value>1)%>%
filter(scale=="Cat")%>%
mutate(year=as.numeric(year),
       value=round(value,1),
       value=format(value, scientific = FALSE))

pal <- colorNumeric("plasma",  domain=unique(MTBS_mapdf1$year))

map1<-leaflet(MTBS_mapdf1) %>%
  addProviderTiles("Stamen.Terrain") %>%
  addCircleMarkers(
    color = ~pal(year),
    stroke = FALSE, fillOpacity = 0.5,
    lng = ~lake_lon_decdeg, lat = ~lake_lat_decdeg,
    label = ~as.character(year),
    popup = ~paste0("Lake name: ",lake_namegnis,
                    "<br/>COMID: ",nhdplusv2_comid,
                   "<br/>Percentage burned: ",value,
                   "<br/>Year: ",year)) %>%
  addLegend("bottomleft",
            title="Year of burn",
            pal = pal, values = ~year,
            labFormat = labelFormat(big.mark="")) 
map1
```

If you zoom in on just california you can see that there is a large group of lakes in northern California that burned in the late 2010s. There is another group of lakes in southern California that burned about 10 years earlier. Will be interesting to see the burn severity on those sites, and look at their recovery trajectory. 


### Maps -- spatial structure of burn severity

This version of the map lets us look at lakes based on burn severity classes. I created this tool so we could identify clusters of high severity burn areas that affected lakes. 

```{r, echo=FALSE, message=FALSE, warning=FALSE,fig.width=8, fig.height=7}
MTBS_mapdf2 <- MTBS_long %>%
                      filter(value>1)%>%
                      group_by(scale)%>%
                      mutate(burnSeverityClass=
                              ifelse(value >1 & value <=20, "< 20%", 
                                     ifelse(value >20 & value <=40, "20-40%", 
                                            ifelse(value >40 & value <=60, "40-60%", 
                                                   ifelse(value >60 & value <=80, "60-80%",
                                                          ifelse(value >80 , "> 80%", "error")))))) %>%
                      mutate(burnSeverityClass = factor(burnSeverityClass,
                                                levels = c("< 20%", "20-40%", "40-60%",
                                                           "60-80%", "> 80%"))) %>% #Reorder factors for plotting
                      filter(scale=="Cat") %>%
  mutate(year=as.numeric(year),
       value=round(value,1),
       value=format(value, scientific = FALSE))

pal <- 
   colorFactor(palette = c("#264653","#2a9d8f","#e9c46a","#f4a261","#e76f51"),
               levels = c("< 20%", "20-40%", "40-60%","60-80%", "> 80%"))


## Toggle the burn severity indices with some layering
## https://community.rstudio.com/t/addlayerscontrol-displays-only-one-base-layer/63899/2
less20<-MTBS_mapdf2 %>% filter(burnSeverityClass=="< 20%")
less40<-MTBS_mapdf2 %>% filter(burnSeverityClass=="20-40%")
less60<-MTBS_mapdf2 %>% filter(burnSeverityClass=="40-60%")
less80<-MTBS_mapdf2 %>% filter(burnSeverityClass=="60-80%")
over80<-MTBS_mapdf2 %>% filter(burnSeverityClass=="> 80%")

leaflet(MTBS_mapdf2) %>%
  addProviderTiles("Stamen.Terrain") %>%
  addCircleMarkers(data = less20, # first group
                   radius = 3.5,
                   fillOpacity = .5,
                   stroke = FALSE,
                       lng = ~lake_lon_decdeg, lat = ~lake_lat_decdeg,
                   popup = ~paste0("Lake name: ",lake_namegnis,
                    "<br/>COMID: ",nhdplusv2_comid,
                   "<br/>Percentage burned: ",value,
                   "<br/>Year: ",year),
                   color = ~pal(burnSeverityClass),
                   # clusterOptions = markerClusterOptions(),
                   group = "< 20%") %>% 
  addCircleMarkers(data = less40, # second group
                   radius = 3.5,
                   fillOpacity = .5,
                   stroke = FALSE,
                       lng = ~lake_lon_decdeg, lat = ~lake_lat_decdeg,
                   popup = ~paste0("Lake name: ",lake_namegnis,
                    "<br/>COMID: ",nhdplusv2_comid,
                   "<br/>Percentage burned: ",value,
                   "<br/>Year: ",year),
                   color = ~pal(burnSeverityClass),
                   # clusterOptions = markerClusterOptions(),
                   group = "20-40%") %>% 
    addCircleMarkers(data = less60, # third group
                   radius = 3.5,
                   fillOpacity = .5,
                   stroke = FALSE,
                       lng = ~lake_lon_decdeg, lat = ~lake_lat_decdeg,
                   popup = ~paste0("Lake name: ",lake_namegnis,
                    "<br/>COMID: ",nhdplusv2_comid,
                   "<br/>Percentage burned: ",value,
                   "<br/>Year: ",year),
                   color = ~pal(burnSeverityClass),
                   # clusterOptions = markerClusterOptions(),
                   group = "40-60%") %>% 
      addCircleMarkers(data = less80, # fourth group
                   radius = 3.5,
                   fillOpacity = .5,
                   stroke = FALSE,
                       lng = ~lake_lon_decdeg, lat = ~lake_lat_decdeg,
                   popup = ~paste0("Lake name: ",lake_namegnis,
                    "<br/>COMID: ",nhdplusv2_comid,
                   "<br/>Percentage burned: ",value,
                   "<br/>Year: ",year),
                   color = ~pal(burnSeverityClass),
                   # clusterOptions = markerClusterOptions(),
                   group = "60-80%") %>% 
        addCircleMarkers(data = over80, # fourth group
                   radius = 3.5,
                   fillOpacity = .5,
                   stroke = FALSE,
                   lng = ~lake_lon_decdeg, lat = ~lake_lat_decdeg,
                   popup = ~paste0("Lake name: ",lake_namegnis,
                    "<br/>COMID: ",nhdplusv2_comid,
                   "<br/>Percentage burned: ",value,
                   "<br/>Year: ",year),
                   color = ~pal(burnSeverityClass),
                   # clusterOptions = markerClusterOptions(),
                   group = "> 80%") %>% 
  addLegend(position = "bottomleft",
            values = ~burnSeverityClass,
            # opacity = .7,
            pal = pal, # palette declared previously
            title = "Burn severity class") %>% 
  addLayersControl(overlayGroups = c("< 20%", "20-40%", "40-60%","60-80%", "> 80%"),
                   options = layersControlOptions(collapsed = FALSE))
```

```{r}
MTBS_long %>%
  filter(value>1)%>%
  group_by(scale)%>%
  mutate(burnSeverityClass=
           ifelse(value >1 & value <=20, "< 20%", 
                  ifelse(value >20 & value <=40, "20-40%", 
                         ifelse(value >40 & value <=60, "40-60%", 
                                ifelse(value >60 & value <=80, "60-80%",
                                       ifelse(value >80 , "> 80%", "error")))))) %>%
  mutate(burnSeverityClass = factor(burnSeverityClass,
                             levels = c("< 20%", "20-40%", "40-60%",
                                        "60-80%", "> 80%"))) %>%#Reorder factors for plotting
  group_by(burnSeverityClass) %>%
  summarize(n=length(unique(nhdplusv2_comid))) %>%
  rename(`Number of lakes`=n,
         `Burn severity class`=burnSeverityClass)
```
